---
title: "flowWorkspace and ggcyto"
author: "Jordi Camps"
date: "2018 M03 28"
output: html_document
---

```{r message=TRUE}
library(flowWorkspace)
library(ggplot2)
library(ggcyto)
```

# Load FlowJo workspace
```{r}
(ws <- openWorkspace("data/04-Apr-2018.wsp"))
```

Access list of samples
```{r}
getSamples(ws)
```

Acces groups
```{r}
getSampleGroups(ws)
```

```{r}
sn <- "data/Sample_TY_1_019.fcs"
getKeywords(ws, sn)
```


# Parsing the workspace
In order to get more from the gating tree we need to parse the XML workspace into R data structures to represent the information therein.
```{r message=FALSE}
(gs <- parseWorkspace(ws, name = "Sample"))
```

```{r}
sampleNames(gs)
```

## Basics on GatingSet
plot gating tree
```{r}
plot(gs)
```

Get nodes of the first path
```{r}
getNodes(gs, path = 1)
```

Get nodes of all paths
```{r}
getNodes(gs, path = "full")
```

Get nodes with shortest defined path
```{r}
nodelist <- getNodes(gs, path = "auto")
nodelist
```

Get gate associated to specific population
```{r}
node <- nodelist[7]
(g <- getGate(gs, node))
```

Retrieve population statistics
```{r}
getPopStats(gs)
```

Plot individual gates
```{r}
plotGate(gs, "Lin-7AAD-")
```

Attach metadata to GatingSet
```{r eval=FALSE, include=FALSE}
d <- data.frame(sample = factor(c(paste("FMO", c("Itga7_PE_Vio770", "Alpl_PE", "Sca1_SB436"), sep = "_"), paste("TNAP-YFP", 1:7, sep = " "))))
pd <- pData(gs)
pd <- cbind(pd, d)
pData(gs) <- pd
pData(gs)
```

Retrieve underlying flow data
```{r}
fs <- getData(gs)
class(fs)
```

# Plotting gated data
## Plot gating tree
```{r}
plot(gs)
plot(gs, "Single Cells")
plot(gs, "Lin-7AAD-")
```

## Plot with ggcyto

```{r}
getData(gs[[1]])
```


Plot with ggplot not recommended
```{r}
fs <- getData(gs[[4]], "Lin-7AAD-")
ggplot(fs, aes(y = `Comp-BV421-A`, x = `Comp-FITC-A`)) +
  geom_hex(bins = 128) +
  axis_x_inverse_trans() +
  axis_y_inverse_trans() #+
  #geom_gate("Lin-Live")
```

Data are stored transformed. When plotting, axes are scales to show raw values. The latest patch on RGlab/ggcyto lets you pass axis_inverse_trans=FALSE to autoplot to show raw values (otherwise you need to build up the plot with ggcyto and pass it in there).

```{r fig.height=10, fig.width=10}
ggcyto(gs, aes(y = `Comp-BV421-A`, x = `Comp-FITC-A`)) +
  geom_hex(bins = 128) +
  geom_gate(c("Alpl+Sca1-", "Alpl+Sca1+", "Alpl-Sca1+")) +
  geom_stats() +
  axis_x_inverse_trans() +
  axis_y_inverse_trans() +
  labs_cyto("marker")
```

```{r}
autoplot(fs, "Comp-FITC-A", "Comp-BV421-A", bins = 64)
```


```{r}
autoplot(gs[[4]], c("Alpl+Sca1-", "Alpl+Sca1+", "Alpl-Sca1+"), bins = 64)
```

```{r}
closeWorkspace(ws)
```


